\mythesistitleNOTSYNPOSIS \tnote[fyi]{Synopsis title is ``Generalization of Consecutive Ones Property of Binary Matrices''}

### 2012-01-31 src/010chap1.tex
- Figure 3.1 for hypergraphs added. Preliminaries section complete. - Characterization of FTPL section moreorless done except section for worked out example -- changed h.isomorphism induced tpl notation to \cp_{\phi} since \cl is confusing because it is the input in most places - Filled up some definitions in Chap 1 - Added asteroidal triples def in chapter 2

\section[Basic preliminaries]{Basic preliminaries} 
\label{sec:basicprelim}

@@

\begin{definition}[\emph{Binary matrix}.]
  \label{def:binmatrix}
  Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
  element, i.e. element at $i$th row and $j$th column. $M$ is a binary
  matrix if each of its element is 0 or \un.  In other words, for all
  $i \in [n]$ and $j \in [m]$, $m_{i,j} \in \{0, \un\}$
\end{definition}

\begin{definition}
  A {\em permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
  x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
  $\lambda$ may be written as a sequence $x_{i_1}x_{i_2}\ldots x_{i_n}$,
  where $i_j \in [n]$ to mean that $\lambda(x_j)= x_{i_j}$.
  This document uses both notations as convenient in context.
\end{definition}

\begin{definition}[\emph{\Cop}]
    \label{def:copmatrix}
  \begin{enumerate}
  \item \Cop
  \item Strong \cop
  \item \Cop order or \cop permutation
  \end{enumerate}
\end{definition}

\figcopmatrix
\tnote[important]{have a figure for set systems also}

\begin{definition}[\emph{\Crop}]
  \label{def:cropmatrix}
\end{definition}

\begin{definition}[\emph{set system}]
    \label{def:setsystem}

    also include overlapping sets.
\end{definition}

\begin{definition}[\emph{Graph, Tree}]
  \label{def:graphtree}

\end{definition}
When refering to a tree as $T$ it could be a reference to the tree
itself, or the vertices of the tree. This will be clear from the
context.

\begin{definition}[\emph{Maximal clique}]
  \label{def:maxclique}    
\end{definition}

>>>>

\tnote[dire]{
\section[Basic preliminaries]{Basic preliminaries} 
\label{sec:basicprelim}

@@

\begin{definition}[Binary matrix]
  \label{def:binmatrix}
  Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
  element, i.e. element at $i$th row and $j$th column. $M$ is a \textbf{binary
  matrix} if each of its element is 0 or \un.  In other words, for all
  $i \in [n]$ and $j \in [m]$, $m_{i,j} \in \{0, \un\}$
\end{definition}

\begin{definition}[Permutation]
  A \textbf{permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
  x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
  $\lambda$ may be written as a sequence $x_{i_1}x_{i_2}\ldots x_{i_n}$,
  where $i_j \in [n]$ to mean that $\lambda(x_j)= x_{i_j}$.
  This document uses both notations as convenient in context.
\end{definition}

\begin{definition}[\Cop]
  \label{def:copmatrix}
  Let $M$ be a binary matrix.
  \begin{itemize}
  \item \textbf{\Cop}
  \item \textbf{Strong \cop}
  \item \textbf{\Cop order} or \cop permutation
  \end{itemize}
\end{definition}

\figcopmatrix

\begin{definition}[\Crop]
  \label{def:cropmatrix}
\end{definition}

\begin{definition}[Set system]
  \label{def:setsystem}

  Let $U$ be a universe with $|U| = n$.  
  \begin{itemize}
  \item The set $\F \subseteq (2^{U} \setminus \emptyset)$ is called a
    \textbf{set system} with universe $U$.
  \item Two sets $S, S' \in \cF$ are said to \textbf{overlap}, denoted
    by $S \overlap S'$, if they have a non-empty intersection and
    neither is contained in the other.
    \[S \overlap S' \text{ \iff\ } S \cap S' \ne \emptyset, S
    \nsubseteq S', S' \nsubseteq S\]
  \end{itemize}
\end{definition}

\begin{definition}[Partial order]
  \label{def:poset}    
  Let $A$ be a set and $R$ be a binary relation $R \subseteq A \times
  A$.

  \textbf{poset}

  Let $X$ be a partially ordered set with $\preccurlyeq$ being the
  partial order on $X$.

  \begin{itemize}
  \item \textbf{hasse diagram}
  \item \textbf{maximal upper bound}  $X_m \in X$ is a \textbf{maximal upper bound} of $X$
  if $\nexists X_q \in X$ such that $X_m \preccurlyeq X_q$.  A maximal
  upper bound on $X$ is denoted by $mub(X)$.

  \item A set of sets form a \textbf{single
      inclusion chain} when the Hasse diagram of their subset relation
    forms a single chain.
  \end{itemize}
\end{definition}

\begin{definition}[Graph, Tree]
  \label{def:graphtree}  
  \tnote[dire]{take from doms}
\end{definition}

When refering to a tree as $T$ it could be a reference to the tree
itself, or the vertices of the tree. This will be clear from the
context.

\begin{definition}[Intersection graph]
  \label{def:intersectiongraph}
  Let $\cF$ be a set system. Then its \textbf{intersection graph}
  $\bI(\cF)$ is a graph \stt its vertex set has a bijection to $\cF$
  and there exists an edge between two vertices \iff their
  corresponding sets in $\cF$ have a non-empty intersection.
\end{definition}

\begin{definition}[Interval graph, Path graph]
  \label{def:pathgraph}

  \begin{itemize}
  \item A graph $G$ is an \textbf{interval graph} if there a set of intervals
    $\cI$ \stt $G$ is isomorphic to the intersection graph of
    $\cI$. \[G \cong \bI(\cI)\]
  \item A graph $G$ is a \textbf{path graph} if there exists a tree $T$ and a
    set of paths from $T$, $\cP$ \stt $G$ is isomorphic to the
    intersection graph of $\cP$. \[G \cong \bI(\cP)\]
  \end{itemize}
\end{definition}
Moreover, a \textbf{chordal graph} is characterized as an intersection
graph of subtrees of a tree.

\begin{definition}[Maximal clique]
  \label{def:maxclique}    
\end{definition}
}

### 2012-01-31 src/020chap2.tex
- Figure 3.1 for hypergraphs added. Preliminaries section complete. - Characterization of FTPL section moreorless done except section for worked out example -- changed h.isomorphism induced tpl notation to \cp_{\phi} since \cl is confusing because it is the input in most places - Filled up some definitions in Chap 1 - Added asteroidal triples def in chapter 2

Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
cannot have {\em asteroidal triples}\footnote{If $G = (V,E)$ is a
  graph, a set of three vertices from $V$ form an {\em asteroidal
    triple} if between any two of them there exists a path in $G$ that
  does not contain any vertex from the closed neighborhood of the
  third vertex.} contained in the corresponding vertex
partition\footnote{The partition corresponds to columns (rows) if its
  half adjacency matrix has \COP columns (rows).}.
Theorem~\ref{th:tuckerforbidden} lists the structures in a
bipartite graph that force one of its vertex partitions to have
asteriodal triples -- in other words, it identifies the subgraphs that
prevent the graph from being convex bipartite.

@@

What we saw so far in this section is checking for the feasibility of
a given interval assignment. ICPIA can also be used to find an
interval assignment when only the set system $\cF$ is given. Part of
the approach is similar to the overlap component idea in \cite{fg65,
  wlh02}. Once the overlap components are factored out, they can be
independently assigned a subinterval to which the sets in the
component need to be mapped. This is simple because the overlap
components are disjoint from each other. Now the subproblem is finding
\COP order for each overlap component. In each component, the set
whose overlaps with the rest of the sets in that component forms
a single inclusion chain\footnote{A set of sets form a {\em single
  inclusion chain} when the Hasse diagram of their subset relation
forms a single chain. In this particular reference, the set of sets is
the set of overlaps.} is chosen first and assigned the
leftmost\footnote{By convention. Rightmost can also be chosen
  analogously to get a different \COP order.} interval. The next
candidate set is chosen from this set's overlapping sets. The interval
assigned is calculated by simple intersection cardinality
means\footnote{Assume that the next set overlaps on one of the
  ``ends'' of the current set and calculate the interval by shifting
  left or right of the current interval and adjusting its size to
  match the next set's cardinality.}. Once the next set is assigned an
interval, the current assigned sets are all checked for ICPIA. This is
a backtracking algorithm -- if at any point ICPIA fails, another
overlapping set is chosen. If all overlapping set attempts fail ICPIA,
a different ``first set'' is chosen. This is a remarkably simple
algorithm due to its obvious simplicity in implementation.

>>>>

Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
cannot have {\em asteroidal triples} (see
Definition~\ref{def:asteroidal}) contained in the corresponding vertex
partition. This partition of the bipartite graph corresponds to
columns (rows) if its half adjacency matrix has \COP columns (rows).
Theorem~\ref{th:tuckerforbidden} lists the structures in a bipartite
graph that force one of its vertex partitions to have asteriodal
triples -- in other words, it identifies the subgraphs that prevent
the graph from being convex bipartite.

\begin{definition}[Asteroidal triple]
  \label{def:asteroidal}
  % If $G = (V,E)$ is a graph, a set of three vertices from $V$ form
  % an {\em asteroidal triple} if between any two of them there exists
  % a path in $G$ that does not contain any vertex from the closed
  % neighborhood of the third vertex.
  An {\em asteroidal triple} in a graph $G$ is a tuple of vertices
  $\{v_1,v_2,v_3\} \subset V$ such that between every pair of vertices
  from this tuple there exists a path that avoids the closed
  neighbourbood of the third vertex.
\end{definition}

@@

What we saw so far in this section is checking for the feasibility of
a given interval assignment. ICPIA can also be used to find an
interval assignment when only the set system $\cF$ is given. Part of
the approach is similar to the overlap component idea in \cite{fg65,
  wlh02}. Once the overlap components are factored out, they can be
independently assigned a subinterval to which the sets in the
component need to be mapped. This is simple because the overlap
components are disjoint from each other. Now the subproblem is finding
\COP order for each overlap component. In each component, the set
whose overlaps with the rest of the sets in that component forms a
single inclusion chain (see Definition~\ref{def:poset}. In this
particular reference, the set of sets is the set of overlaps.)  is
chosen first and assigned the leftmost\footnote{By
  convention. Rightmost can also be chosen analogously to get a
  different \COP order.} interval. The next candidate set is chosen
from this set's overlapping sets. The interval assigned is calculated
by simple intersection cardinality means\footnote{Assume that the next
  set overlaps on one of the ``ends'' of the current set and calculate
  the interval by shifting left or right of the current interval and
  adjusting its size to match the next set's cardinality.}. Once the
next set is assigned an interval, the current assigned sets are all
checked for ICPIA. This is a backtracking algorithm -- if at any point
ICPIA fails, another overlapping set is chosen. If all overlapping set
attempts fail ICPIA, a different ``first set'' is chosen. This is a
remarkably simple algorithm due to its obvious simplicity in
implementation.
